% !TEX root = macfp_2017_gasphase.tex

\subsection{Objectives} \label{sec:CPS_session_1}

The production of combustible gases by burning materials is typically the rate-limiting process in the growth of fire.  A quantitative understanding of this process is therefore essential for advancing our ability to predict and mitigate fire development.  Unfortunately, measurement and modeling efforts carried out in this field by various research groups tend to be poorly coordinated.  Little agreement exists as to what constitutes best practices and standards in data collection and model development.  The purpose of the condensed phase subgroup of the MaCFP Working Group is to facilitate data and model sharing among researchers in order to improve predictions of thermal decomposition and pyrolysis in fire.  The work of this subgroup, in conjunction with the work of the gas phase subgroup, is expected to lead to fundamental progress in fire modeling.  It is envisioned that the two subgroups will collaborate to make quantitative predictions of the combined gas-solid phase processes that determine flame spread.  It is also envisioned that the two subgroups will hold joint workshops every two or three years.

The condensed phase subgroup shares the central objective of MaCFP ``to target fundamental progress in fire science and to advance predictive fire modeling."  The specific objectives of the subgroup will focus on the development, calibration, verification, and validation of predictive models of thermal decomposition and pyrolysis.  To this end, the subgroup plans to:

\begin{itemize}
\item Develop several alternative formats for experimental data sets that carry sufficient information to enable parameterization of pyrolysis models for a given material.
\item Develop a set of requirements for data set quality and completeness and organize a committee of experts that will review the submissions to the repository to ensure that they are compliant with these requirements.
\item Incorporate compliant data sets into the existing MaCFP data repository~\cite{MaCFP_repository}.
\item Create a database of pyrolysis property sets that are generated from the experimental data sets.  Each pyrolysis property set will be required to be accompanied by a demonstration of how well it captures the data on the basis of which it was calibrated and validated.
\item Develop a set of minimum requirements for numerical pyrolysis simulation codes.
\item Organize a discussion group focused on unresolved issues in pyrolysis modeling.
\end{itemize}

The scientific topics covered by the condensed phase subgroup will include:

\begin{itemize}
\item Kinetics and thermodynamics of the condensed phase decomposition reactions.
\item Properties and composition of gaseous pyrolyzates.
\item Heat and mass transfer in the condensed phase.
\item Physics and chemistry of the gas-condensed phase interface including the topics of oxidative pyrolysis and interactions with the surface flame.
\item Coupled thermal and mechanical behavior of pyrolyzing solids including intumescence and melt flow.
\end{itemize}

